TODO: A Survey of StarCraft AI Techniques

### 2012-05-11 survey.tex
Clarified that diagrams come from analyzing sourcecode
reworked the first part of the state of the art, intro to problems

\begin{figure*}
    \centering
    \includegraphics[width=0.9\columnwidth]{figures/levels_abstraction.pdf}
    \caption{RTS AI levels of abstraction and their properties: uncertainty (from intentions and from partial observation) is going higher as the abstraction levels are raising. The timings on the right correspond to an estimate of the duration of a behavior switch in StarCraft.}
    \label{fig:levels-abstraction}
\end{figure*}

%\subsection{Integration of Domain Knowledge}\label{subsec:domainknowledge}
{\color{blue}
All bots attempt at integrating human domain knowledge in a way or another, be it in the form of build-orders, heuristics, strategies, or by automatically learning from human replays. This has been shown to be key to good performance in bots (for now, as Dave was saying: ``Starcraft AI is like Chess was in the 60s'').

Talk about learning, scripting, and other techniques to incorporate such knowledge.

When talking about learning in RTS games, with regard to a particular match against an opponent, one has to differentiate between:
\begin{itemize}
\item prior learning: done before the match, like when a bot learns what is likely to happen on a given match-up (and map) from replays, or when a bot optimizes its micro-management parameters through simulated annealing or genetic evolution.
\item in-match learning: done between the different games of match, ``the opponent rushed me last game so I will play this one safe''.
\item in-game learning: done between the different actions of the game, ``I tried to attack here and it was a failure, let us try somewhere else''.
\end{itemize}
All these learning can be supervised (labeled replays, scripted situations) or unsupervised (unlabeled data, exploratory situations). The last two kinds of situations of learning are more inclined towards reinforcement learning because the AI has to deal with the exploration-exploitation duality. %prior learning: supervised \cite{WeberCIG09} \cite{SynnaeveCIG11}, unsupervised: \cite{SynnaeveAIIDE11}
}

%\subsection{Reasoning with Uncertainty}\label{subsec:uncertainty}
{\color{blue}
I think there are two important aspects of uncertainty in RTS games:
\begin{itemize}
\item First is trying to model uncertainty and try to reduce it (information gathering, and representation of knowledge), this includes scouting, and maintaining good estimates of locations and number of units of the enemy, etc.
\item Second is actually using the uncertain information. Taking decisions like army strength estimation (like to attack or retreat).
\end{itemize}

There are two main kinds of uncertainty, existing for two different reasons, in RTS games. First, there is the incompleteness of information, due to the fog of war, which leads to some kind of ``extensional'' 
%``extensive'' 
uncertainty, that can be lowered by good scouting, and knowledge representation and reasoning (to infer what is possible given what has been seen). Also, there is the fact that we will never be able to read the opponent's mind, the ``intentional'' 
%or ``intensive'' 
uncertainty that we can't really be sure to scratch off, even with the best scouting possible. For intentional uncertainty, the AI, as the human player, can only build a sensible model of what the opponent is likely to do, to think, or to think knowing what we think he thinks. Both these kinds of uncertainty can be studied through probabilities, but only $P(Hidden | Seen)$ can be \textit{directly} linked to data only for extensional uncertainty. %now for the clumsy part:
One could argue that the extensional data is the only thing that matters because it is the only thing that has an influence on the game, but in reality, for very highly skilled players, discovering the intentions of the enemy is the key to winning. It allows for a compact understanding of the game state and development, while also giving additional information about the future. Whereas a player is doing action A to follow it up by action B or C depends on its intention, while the information about action A is visible and information about B or C is hidden, only intention allows for a choice.
}

@@

\begin{figure*}[ta]
    \centering
    \includegraphics[width=\textwidth]{figures/figure-bot-architectures-wide.pdf}
    \caption{Architecture of 6 Starcraft bots. Modules with black background sent commands directly to Starcraft, dashed arrows represent data flow, and solid arrows represent control.}
    \label{fig:bot-architecture}
\end{figure*}

>>>>

\subsection{Problems}

{\color{blue}
\textbf{Planning}
TODO (Santi?) just a paragraph on exposing why there is a planning problem in RTS AI.

\textbf{Spatial and Temporal Reasoning}
TODO just a paragraph on exposing why there are temporal and spatial reasoning in RTS AI.

\textbf{Learning}\\
When talking about learning in RTS games, with regard to a particular match against an opponent, one has to differentiate between:
\begin{itemize}
\item prior learning: done before the match, like when a bot learns what is likely to happen on a given match-up (and map) from replays, or when a bot optimizes its parameters beforehand through simulated annealing or genetic evolution.
\item in-game learning: done between the different actions (decisions) taken during a game. For instance, a bot can take into account the successes and failures of its attacks (location, type, timings).
\item in-match learning: done between the different games of match, it encompasses opponent modeling across multiple games and adapting the bots' strategy to its successes and failures.
\end{itemize}
All these learning can be supervised (labeled replays, scripted situations) or unsupervised (unlabeled data, exploratory situations). The last two kinds of situations of learning are more inclined towards reinforcement learning because the AI has to deal with the exploration-exploitation duality. %prior learning: supervised \cite{WeberCIG09} \cite{SynnaeveCIG11}, unsupervised: \cite{SynnaeveAIIDE11}

%%% I think there are two important aspects of uncertainty in RTS games:
%%% \begin{itemize}
%%% \item First is trying to model uncertainty and try to reduce it (information gathering, and representation of knowledge), this includes scouting, and maintaining good estimates of locations and number of units of the enemy, etc.
%%% \item Second is actually using the uncertain information. Taking decisions like army strength estimation (like to attack or retreat).
%%% \end{itemize}

\textbf{Uncertainty}\\
In RTS games, there are two main kinds of uncertainty, existing for two different reasons. First, there is incompleteness of information, due to the fog of war, which leads to some kind of ``extensional'' 
%``extensive'' 
uncertainty, that can be lowered by good scouting, and knowledge representation (to infer what is possible given what has been seen). Secondly, there is the fact that we will never be able to read the opponent's mind, the ``intentional'' 
%or ``intensive'' 
uncertainty that we cannot really be sure to scratch off, even with the best scouting possible. For intentional uncertainty, the AI, as the human player, can only build a sensible model of what the opponent is likely to do, to think, or to think knowing what we think he thinks. Both these kinds of uncertainty can be studied through probabilities, but only for extensional uncertainty can $P(Hidden | Seen)$ be \textit{directly} linked to data. %now for the clumsy part:
One could argue that the extensional data is the only thing that matters because it is the only thing that has an influence on the game, but in reality, for very highly skilled players, discovering the intentions of the enemy is the key to winning. It allows for a compact understanding of the game state and development, while also giving additional information about the future. Whereas a player is doing action A to follow it up by action B or C depends on its intention, while the information about action A is visible and information about B or C is hidden, only intention allows for a choice.
}

\subsection{Decomposition}
\begin{figure*}
    \centering
    \includegraphics[width=0.9\columnwidth]{figures/levels_abstraction.pdf}
    \caption{RTS AI levels of abstraction and their properties: uncertainty (from intentions and from partial observation) is going higher as the abstraction levels are raising. The timings on the right correspond to an estimate of the duration of a behavior switch in StarCraft. Spatial and temporal reasoning are indicated for part for which greedy solutions are really not efficient.}
    \label{fig:levels-abstraction}
\end{figure*}
As a work on RTS AI has to address several of these aforementioned problems together, we decided to study decompose RTS AI in three tasks, widely known by RTS players: strategy, tactics and units control (micro-management), see Figure~\ref{fig:levels-abstraction}. We consider strategy to be everything related to the \textit{technology tree}, being the buildings and the upgrades, and the army composition. It is the most intentional level, as a player performs a strategy with future stances (aggressive, defensive, economy, technology) and tactics in mind. To change its strategy, we consider that a player need about 3 minutes: they should need to construct a few new buildings and/or research some upgrades and/or produce units changing the army composition. Tactics represent the attacks and defenses. For the human player, decisions on tactics are made from spatial (exploiting the terrain) and temporal (army movements) reasoning, constrained on the possible types of attacks by the army composition of the player and their opponent. Units control, also called micro-management, is how the player takes action to maximize the efficiency of its units (applying the tactics) in real-time (simultaneous) and adversary conditions. As shown in Figure~\ref{fig:levels-abstraction}, strategy constrains possible tactics which in turn constrain units control. On a prediction/adaptation point of view, the units observation level gives information about the opponent's tactics and strategy. Flaws in opponents tactics can also steer our strategy into exploiting them.\\
perhaps TODO examples?

@@

\begin{figure*}[ta]
    \centering
    \includegraphics[width=\textwidth]{figures/figure-bot-architectures-wide.pdf}
    \caption{Architecture of 6 Starcraft bots obtained by analyzing their source code. Modules with black background sent commands directly to Starcraft, dashed arrows represent data flow, and solid arrows represent control.}
    \label{fig:bot-architecture}
\end{figure*}
